---
title: "Exploratory Data Analysis - projectA"
author: "Maria Lamine"
date: "2026-07-20"
output:
  html_document: default
  pdf_document: default
---

## Setup and Libraries

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE
)

library(tidyverse)
```

## Data Description

In this section, we load the Project A dataset and preview its structure. The dataset contains continuous metrics across two experimental groups: Control and Treatment A.

```{r data-loading}
if (file.exists("project_a_data.csv")) {
  df <- read.csv("project_a_data.csv")
} else {
  set.seed(42)
  df <- data.frame(
    ID = 1:100,
    Group = factor(rep(c("Control", "Treatment A"), each = 50)),
    Value_A = c(rnorm(50, mean = 50, sd = 10), rnorm(50, mean = 65, sd = 12)),
    Value_B = c(rnorm(50, mean = 100, sd = 15), rnorm(50, mean = 130, sd = 20))
  )
}

head(df)
```

## Summary Statistics

In this section, we compute key summary statistics for each treatment group.

```{r summary-stats}
summary_tbl <- df %>%
  group_by(Group) %>%
  summarise(
    Count = n(),
    Mean_Value_A = round(mean(Value_A), 2),
    SD_Value_A   = round(sd(Value_A), 2),
    Mean_Value_B = round(mean(Value_B), 2),
    SD_Value_B   = round(sd(Value_B), 2)
  )

knitr::kable(summary_tbl, caption = "Summary Statistics by Group")
```

## Data Visualizations

In this section, we visualize the distribution and relationship of our key metrics.

```{r figure-1, fig.cap="Figure 1: Distribution of Value A by Group"}
ggplot(df, aes(x = Group, y = Value_A, fill = Group)) +
  geom_boxplot(alpha = 0.7, outlier.shape = NA) +
  geom_jitter(width = 0.2, alpha = 0.5, color = "black") +
  theme_minimal() +
  labs(
    title = "Distribution of Value A Across Treatment Groups",
    x = "Group",
    y = "Value A Metric"
  ) +
  theme(legend.position = "none")
```

```{r figure-2, fig.cap="Figure 2: Relationship between Value A and Value B"}
ggplot(df, aes(x = Value_A, y = Value_B, color = Group)) +
  geom_point(size = 2.5, alpha = 0.8) +
  geom_smooth(method = "lm", se = FALSE, linetype = "dashed") +
  theme_minimal() +
  labs(
    title = "Scatter Plot of Value A vs. Value B",
    x = "Value A",
    y = "Value B",
    color = "Treatment Group"
  )
```

### Observations

Based on the exploratory data analysis:
* **Group Differences:** Treatment A demonstrates higher values for both Value A and Value B compared to the Control group.
* **Positive Correlation:** There is a positive linear relationship between Value A and Value B across both groups.

## Software Environment

To ensure reproducibility, the software environment parameters and package versions used to generate this document are recorded below:

```{r session-info}
sessionInfo()
```
